---
title: "Week 8, Day 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(PPBDS.data)
library(rstanarm)
library(tidyverse)
library(rsample)

# Recall that Enos (2014) actually looks at a measure of change in attitude
# toward immigration, rather than simply at the ending attitude. Let's create
# such a variable and try to model it.

week_8 <- trains %>% 
  mutate(att_chg = att_end - att_start) %>% 
  select(att_chg, gender, liberal, party, treatment)
  
```

Weeks 7, 8 and 9 are the core of the class. We have a question. We have some data. How should we use the data to answer the question? Using Wisdom, we first decide if the question and the data are "close enough" that we can consider them to both be part of the same population. With Justice, we create a mathematical model which describes the connection between the outcome we want to explain/understand and the covariates which might be connected to it. Courage takes us from mathematics to code, creating a model, including posterior distributions for all its parameters. The last step is to use that model to answer the question with which we started, with Temperance.




## Scene 1

**Prompt:** Let's estimate and interpret three models. In each case, `att_chg` is the outcome variable, the thing we are trying to understand and explain. Use `stan_glm()` to create and print a model with `att_chg` (the dependent or outcome variable) as a function of the indicated independent (or predictor) variable. Do not use an intercept. In each case, write three sentences of interpretation, with a special focus on whether the model is predictive or causal. For causal models, indicate how you might estimate the causal effect.

* `att_chg` as a function of `gender`


* `att_chg` as a function of `liberal`


* `att_chg` as a function of `treatment`

```{r}

# The first model assesses the effect based on gender. This is a predictive
# model as this variable cannot be changed for each observation. The data shows
# that differences in the results by gender leans toward positive for Female, or
# that the median observation for females will shift towards a more conservative
# outcome, while the median Male observation shows a more liberal outcome.
# The model indicates that men are more likely to have a more liberal view after
# the experiment than women.

stan_glm(att_chg ~ gender - 1,
         data = week_8,
         refresh = 0)

# This model is also predictive rather than causal. The liberal/not liberal 
# variable simply indicates an existing factor that cannot be tested or changed
# in the experiment. The variation between both groups is very small, at 0.1
# difference in the median of both groups.

stan_glm(att_chg ~ liberal - 1,
         data = week_8,
         refresh = 0)

# The last model is the only causal model. Treatment is not an inherent trait 
# of the observations but rather a change that can cause the difference in 
# outcome. If you calculated the average att_change by treatment, you could
# come up with an estimate of the causal effect. The treated observations had a
# median change in attitude of 0.4, or a more conservative attitude towards 
# immigrants after the experiment. In contrast, observations without exposure
# to immigrants during the experiment had a more liberal change in attitude
# of about -0.4 magnitude (which could be explained by omitting variables, etc).

q1 <- stan_glm(att_chg ~ treatment + liberal - 1,
         data = week_8,
         refresh = 0)
```


## Scene 2

**Prompt:** For the model with `att_chg` as a function of `treatment`, create posterior probability densities for the two important parameters. Write a paragraph interpreting the meaning of those posteriors. 

```{r}

# The model shows us that attitude changes differ largely based on whether the
# observation is treated or not. Observations without exposure (control) are
# more likely to have a negative attitude change (towards liberal views), while
# treated observations are more likely to have a positive attitude change
# (towards conservative views). Both distributions seem to be grouped at a 
# similar distance of 0.5 change on average from no change (0).

q1 %>%
  as_tibble() %>%
  select(-sigma) %>%
  mutate(Treated = treatmentTreated, 
         Control = treatmentControl) %>%
  pivot_longer(cols = Treated:Control,
               names_to = "parameter",
                 values_to = "attitude_change") %>%
  ggplot(aes(x = attitude_change, fill = parameter)) +
  geom_density(alpha = 0.5)
#  geom_histogram(aes(y = after_stat(count/sum(count))),              
     #            alpha = 0.75, bins = 100) +
scale_y_continuous(labels=scales::percent_format()) +
theme_classic() +
  labs(x = "Attitude Change",
       y = "Probability",
       title = "Change in Attitude Based on Exposure to Immigrants ")

```



## Scene 3

**Prompt:** Create a plot of the the posterior for the average treatment effect. Write a paragraph interpreting the plot.

```{r}

# The average treatment effect of running the trains experiment is shown in this
# distribution model to be around a 0.5 to 1.0 positive shift in attitude. This
# means that exposure to immigrants on average creates a more conservative view
# of immigration by a small degree.
# Our vline shows the median value of this distribution of average treatment 
# effects at around 0.7944881, a significantly positive change in attitude
# when exposed to treatment!

q <- q1 %>%
  as_tibble() %>%
  select(-sigma) %>%
  mutate(ATE = treatmentTreated - treatmentControl)

  ggplot(q, aes(x = ATE, y = after_stat(count/sum(count)))) +
    geom_density(alpha = 0.5, fill = "dodgerblue") +
 # geom_histogram(alpha = 1, bins = 100, fill = "dodgerblue") +
scale_y_continuous(labels=scales::percent_format()) +
theme_classic() +
  labs(x = "Treatment Effect",
       y = "Probability",
       title = "Average Treatment Effect of Trains Experiment") +
  geom_vline(xintercept = median(q$ATE),
             color = "red",
             lty = "dashed")

  median(q$ATE)
```


```{r}

d <- trains %>%
  filter(treatment == "Treated") %>%
  select(att_end) %>%
  bootstraps(times = 1000) %>%
  mutate(boot = map(splits, ~ analysis(.))) %>%
  mutate(att_end = map(boot, ~ pull(., att_end))) %>%
mutate(att_end_mean = map_dbl(att_end, ~mean(.)))
mean(d$att_end_mean)
```
```{r}

model <- stan_glm(att_chg ~ 1,
         data = week_8,
         refresh = 0)

model_interval <- posterior_interval(model)
model_interval


```

```{r}

model2 <- stan_glm(att_chg ~ treatment,
         data = week_8,
         refresh = 0)
model2_interval <- posterior_interval(model2)

```

